Clustering Beat-Chroma Patterns in a Large Music Database

### 2010-03-23 Paper/Paper.tex
improved

Note that we do not claim that any of these informations (segments, beats, bars)
are perfectly accurate (\eg in \cite{Barrington2009a} authors obtain a better
song segmentation than Echonest). Practise showed us that they are reasonable, 
and the size of the data set should make up for the imperfections or noise.
We also believe that patches sizes based on a number of beats or bars are more
meaningful than an arbitrary time length. More on this in the experiments
(Section \ref{sec:experiments} TO BE CHECKED).

@@

\subsection{Pattern Creation}
Patterns are defined as a number of beats or bars. In the case of beats,
we simply cut the set of chromas per beat every $N$ beats. In the case
of bars, we gather all beats corresponding to the $N$ bars requested.
The number of chroma vectors we get can vary. We resize the features in the
time axis to a fixed length. Resizing can be argued over, but it should
preserve the main bar characteristics (such as a strong beat at the beginning
or a note change in the middle).

@@

\subsection{Testing: USPOP}\label{sec:testdata}
We had access to a low quality (32kbps) version of the songs in the uspop 2002 
data set \cite{uspop2002}.
This well-known data show a great diversity of pop songs, but is not particular
otherwise\footnote{We 
also tried to use Tzanetakis dataset \cite{Tzanetakis2002a}, but 
the $30$ seconds segments seemed hard to analyze by the Echo Nest API, most did
not contain any bar.}.
uspop2002 serves as testing set to measure how well a codebook learned on
the Cowbell data set can represent new songs. We get Echo Nest features
for $8651$ songs of that dataset.

\subsection{Setting}\label{ssec:setting}
We take one or two bars, normalize the patches to size 4, 8, or 16.
We roll the patch to be invariant to the key (on the patch level, not on
the song level). We learn a codebook of size $K$ over the cowbell dataset 
using the VQ algorithm (Algorithm \ref{algo:vq}). A typical learning rate 
is $0.01$ for $200$ iterations over the whole dataset.

@@

% NEW RESULTS COMING SOON WITH FIXED SIZED DATASET
\begin{table}
\begin{center}
\begin{tabular}{|l|l|c|}
\hline
\# codes & distortion \\ \hline \hline
1 & $0.066081$ \\
10 & $0.045579$ \\
50 & $0.038302$ \\
100 & $0.035904$ \\
500 & $0.030841$ \\ \hline
\end{tabular}
\end{center}
\caption{\small{Experiment with the codebook size. We train on patterns
corresponding to $1$ bar resized to $4$ beats. 
The number of training samples is $50000$.
}}
\label{tab:cbsize}
\end{table}

@@

\subsection{Song Segmentation}
We explained that we used bars as defined by the EchoNest algorithm to
train our patterns. Let's assume once again that those bars are
reasonable. We investigate whether we could recover the bar segmentation
of new songs from our codebook. It would helps us assert that codes
encodes something related to bars, for instance a strong beat at the
beginning. Results are shown \ref{tab:offset}. We find the right segment
with an accuracy of $62\%$ where random is $25\%$. Details of the
experiment follow.

>>>>

Note that we do not claim that any of these informations (segments, beats, bars)
are perfectly accurate (\eg in \cite{Barrington2009a} authors obtain a better
song segmentation than Echonest). Practise showed us that they are reasonable, 
and the size of the data set should make up for the imperfections or noise.
We also believe that patches sizes based on a number of beats or bars are more
meaningful than an arbitrary time length. More on this in the experiments
(Subsection \ref{ssec:segment}).

@@

\subsection{Pattern Creation}
Patterns are defined as a number of beats or bars. In the case of beats,
we simply cut the set of chromas per beat every $N$ beats. In the case
of bars, we gather all beats corresponding to the $N$ bars requested.
The number of chroma vectors we get can vary. We resize the features in the
time axis to a fixed length. Resizing over clipping can be argued over, 
but it should
preserve the main bar characteristics (such as a strong beat at the beginning
or a note change in the middle).

@@

\subsection{Testing: USPOP}\label{sec:testdata}
We had access to a low quality (32kbps) version of the songs in the uspop 2002 
data set \cite{uspop2002}.
This well-known data show a great diversity of pop songs, but is not particular
otherwise\footnote{We 
also tried to use Tzanetakis dataset, 
%\cite{Tzanetakis2002a},
but the $30$ seconds segments seemed hard to analyze by the Echo Nest API, 
most did not contain any bar.}.
uspop2002 serves as testing set to measure how well a codebook learned on
the Cowbell data set can represent new songs. We get Echo Nest features
for $8651$ songs of that dataset.

\subsection{Setting}\label{ssec:setting}
We take one or two bars and normalize the patches to size 4 or 8 respectively.
We roll the patch to be invariant to the key (on the patch level, not on
the song level). We learn a codebook of size $K$ over the cowbell dataset 
using the VQ algorithm (Algorithm \ref{algo:vq}). A typical learning rate 
is $0.01$ for $200$ iterations over the whole dataset.

@@

\begin{table}
\begin{center}
\begin{tabular}{|l|l|c|}
\hline
\# codes & distortion \\ \hline \hline
1 & $0.066081$ \\
10 & $0.045579$ \\
50 & $0.038302$ \\
100 & $0.035904$ \\
500 & $0.030841$ \\ \hline
\end{tabular}
\end{center}
\caption{\small{Experiment with the codebook size. We train on patterns
corresponding to $1$ bar resized to $4$ beats. 
The number of training samples is $50000$.
}}
\label{tab:cbsize}
\end{table}

@@

\subsection{Song Segmentation} \label{ssec:segment}
We explained that we used bars as defined by the EchoNest algorithm to
train our patterns. Let's assume once again that those bars are
reasonable. We investigate whether we could recover the bar segmentation
of new songs from our codebook. It would helps us assert that codes
encodes something related to bars, for instance a strong beat at the
beginning. Results are shown \ref{tab:offset}. We find the right segment
with an accuracy of $62\%$ where random is $25\%$. Details of the
experiment follow.
